Finch: Sparse and Structured Array Programming with Control Flow

### 2024-03-21 sections/intro.tex
Update on Overleaf.

Sparse arrays (which store only nonzero elements) describe networks and
simulations.
%
Run-length encoding describes images and masks, geometry, and
databases (such as a list of transactions with the date field all the same)
Symmetry, bands, padding, and blocks are all natural matrix properties which
arise in the course of scientific computing or simulations of the natural world.
%

Operators, such as convolution, can be expressed as structured arrays.
%
For
example, a convolution with a filter can be expressed as a matrix multiplication
with the toeplitz matrix of all the circular shifts of the filter.

\help{This section needs citations}
\textbf{Currently, support for structured data is fragmented and incomplete}.
%
Experts must hand write variations of even the simplest kernels, like matrix
multiply, for each data structure and architecture to get performance.
%
Implementations must choose a small set of features to support well, resulting
in a compromise between \textbf{program flexibility} and \textbf{datastructure
flexibility}.
%
Hand-written solutions are collected in diverse libraries like
MKL, OpenCV, or SciPy. 
%
However, libraries will only ever support a subset of
programs on a subset of datastructure combinations. Even the most advanced
libraries, such as the GraphBLAS, which support a wide variety of sparse
operations over various semirings always lack support for other features, such
as tensors, fused outputs, or runs of repeated values. 
%
While dense array
compilers support an enormous variety of program constructs like early break and
multiple left hand sides, they only support dense arrays.  
%
Special-purpose
compilers like TACO or CoRa which support a select subset of structured data
structures (only sparse, or only ragged arrays) must compromise by greatly
constraining the classes of programs which they support, such as tensor
contractions \cite{kjolstad_tensor_2019, fegade_cora_2022}. 
%
This trade-off is
visualized in figure \ref{fig:datavsprogram}.

>>>>

Sparse arrays (which store only nonzero elements) describe networks and
simulations.
%
Run-length encoding describes images and masks, geometry, and
databases (such as a list of transactions with the date field all the same).
%
Symmetry, bands, padding, and blocks arise due to modeling choices in scientific computing as well as in intermediate structures in many linear solvers.
%
Operators, such as convolution, can be expressed as structured arrays.
%
For example, a convolution with a filter can be expressed as a matrix multiplication
with the toeplitz matrix of all the circular shifts of the filter~\cite{sze2017efficient}.
%
\help{This section needs citations}
\textbf{Currently, support for structured data is fragmented and incomplete}.
%
Experts must hand write variations of even the simplest kernels, like matrix
multiply, for each data structure and architecture to get performance.
%
Implementations must choose a small set of features to support well, resulting
in a compromise between \textbf{program flexibility} and \textbf{datastructure
flexibility}.
%
Hand-written solutions are collected in diverse libraries like
MKL, OpenCV, or SciPy. 
%
However, libraries will only ever support a subset of
programs on a subset of datastructure combinations. Even the most advanced
libraries, such as the GraphBLAS, which support a wide variety of sparse
operations over various semirings always lack support for other features, such
as tensors, fused outputs, or runs of repeated values. 
%
While dense array
compilers support an enormous variety of program constructs like early break and
multiple left hand sides, they only support dense arrays.  
%
Special-purpose
compilers like TACO or CoRa which support a select subset of structured data
structures (only sparse, or only ragged arrays) must compromise by greatly
constraining the classes of programs which they support, such as tensor
contractions \cite{kjolstad_tensor_2019, fegade_cora_2022}. 
%
This trade-off is
visualized in figure \ref{fig:datavsprogram}.
